---
title: "Betke-Briana-ADA-homework-3"
author: "Briana Betke"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 3

Let's start with reading in the appropriate packages for the assignment:
```{r  message=FALSE}
library(tidyverse)
library(broom)
library(cowplot)
theme_set(theme_cowplot())
```
### Challenge 1:

#### Step 1:
Read in the Kamilar and Cooper dataset from online:
```{r message=FALSE}
# reading data
f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/KamilarAndCooperData.csv"
d <- read_csv(f, col_names = TRUE)
head(d,6) # just the first 6 so its not so much when knitted.
```

#### Fit the regression models
Starting with untransformed:
```{r}
mod <- lm(MaxLongevity_m~Brain_Size_Species_Mean, data = d)
tidy(mod) # model output, but nice
```

Now the transformed one:
```{r}
log_mod <- lm(log(MaxLongevity_m)~log(Brain_Size_Species_Mean), data = d)
tidy(log_mod)
```

#### Superimpose line on ggplot and add model equation to the plot
First, I wanted to take out the coefficients of the models to embed into the equation for the graph. So I save the tidy model output as a data frame then put them into the paste function to test how it would look. This took a while because I am bad with paste functions.....apparently. 
```{r}
values <- as.data.frame(tidy(mod)) # the data frame

# testing the paste
paste("longevity = ", paste(round(values$estimate[1],2), paste(round(values$estimate[2],2), "brain size", sep = "*"), sep=" + "), sep="")
```

transformed version
```{r}
log_values <- as.data.frame(tidy(log_mod))

paste("longevity = ", paste(round(log_values$estimate[1],2), paste(round(log_values$estimate[2],2), "brain size", sep = "*"), sep=" + "), sep="")
```

Now on to ggplot! Adding the equation to the graph using the geom_text().
```{r message=FALSE, warning=FALSE}
mod_plot <- ggplot(d, aes(y = MaxLongevity_m, x = Brain_Size_Species_Mean)) +
        geom_point()+
        geom_smooth(method = "lm", color = "black", formula = y ~ x) + 
        geom_text(x = 170, y = 950, size = 3, label = paste("longevity = ",
                                                  paste(round(values$estimate[1],2),
                                                        paste(round(values$estimate[2],2),
                                                              "brain size", sep = "*"), sep=" + "), sep=""))

log_mod_plot <- ggplot(d, aes(y = log(MaxLongevity_m), x = log(Brain_Size_Species_Mean))) +
        geom_point()+
        geom_smooth(method = "lm", color = "black", formula = y ~ x) + 
        geom_text(x = 2.5, y = 6.60, size = 3, label = paste("longevity = ",
                                                  paste(round(log_values$estimate[1],2),
                                                        paste(round(log_values$estimate[2],2),
                                                              "brain size", sep = "*"), sep=" + "), sep=""))
# multiplot time 
library(patchwork)
mod_plot + log_mod_plot
```

#### Hypothesis test

H0: There is no linear relationship (β1=0)
HA: There is a linear relationship (β1≠0)

The p-value for the explanatory variable is less than 0.05 and even siginificant at 0.10. Indicating that there is significant evidence of a linear relationship. 

Viewing the model again, the estimate is 1.21. This can be interpreted as the increase in longevity for every 1 unit increase in mean brain size is 1.21, on average. 
```{r}
values
log_values
```

90% CI for slope estimate untransformed:
```{r}
alpha <- 0.10
(CI <- tidy(mod, conf.int = TRUE, conf.level = 1 - alpha))
```

90% CI for slope estimate untransformed:

#### Add 90% CI and Prediction interval to plot
```{r}

```

#### Point estimate
```{r}


```


#### Looking @ the models, which is better and why?
looking at dist of variables and residuals, make your case bro 

scale_color_hue(labels = c("Fit", "CI", "PI")) 

### Challenge 2:
```{r}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
